# Reproducible Research: Peer Assessment 1

Load required libraries
``` {r LibLoad}
library(plyr)
library(ggplot2)
```

## Loading and preprocessing the data

Reading the data from activity.csv.

```{r LoadData ,  cache=TRUE}
actData= read.csv("./activity.csv",na.strings=c("NA", "NULL"))

```
Structure of the data looks like...

```{r SummaryOfData }
str(actData)

```

Data with ignoring mising values will be:
``` {r MissingValue}
subactData=subset(actData,! is.na(actData$steps))

str(subactData)
```


## What is mean total number of steps taken per day?

Please Note:Missing values in the dataset are being ignored .
``` {r MeanValue}

datestepsdata=tapply(subactData$steps,subactData$date,sum)
dfdatestepsdata=as.data.frame(datestepsdata)
dfdatestepsdata[,2]=unique(actData$date)
names(dfdatestepsdata)=c("stepsmean","date")

qplot(dfdatestepsdata[,1], xlab = expression(bold("Total number of Steps")),  geom = "bar",binwidth=1000)


meansteps=mean(dfdatestepsdata[,1],na.rm=TRUE)
medsteps=median(dfdatestepsdata[,1],na.rm=TRUE)
```
Mean of total number of steps taken per day is `r meansteps`  
Median of total number of steps taken per day is `r medsteps`


## What is the average daily activity pattern?
Please Note:Missing values in the dataset are being ignored .

Time Series of 5-minute interval 
``` {r AvgDailyValue}
intstepsdata=tapply(subactData$steps,subactData$interval,mean)
dfintstepsdata=as.data.frame(intstepsdata)
dfintstepsdata[,2]=unique(actData$interval)
names(dfintstepsdata)=c("avgdailysteps","interval")

ggplot(dfintstepsdata, aes( interval,avgdailysteps)) + geom_line() +    xlab("5-minute Interval") + ylab("Average number of steps taken, averaged across all days")


```

```{r Maxsteps}
maxsteps=max(dfintstepsdata$avgdailysteps)
intmaxsteps=dfintstepsdata[dfintstepsdata$avgdailysteps==maxsteps,]
```

5-minute Interval with maximum number of steps is `r intmaxsteps$interval`


## Imputing missing values

1. Calculating total number f missin values in the dataset

``` {r missingvalcount}
nacount=dim(actData[is.na(actData$steps),])[1]
```

Total number of rows with NA s are: `r nacount`

============================
2. Stratergy for filling in all of themissing Values in the dataset

We plan to impute missing values with mean of number steps of all observations

``` {r imputingNA}
impactData=actData
impactData[is.na(impactData$steps),]$steps=rep(mean(impactData$steps,na.rm=TRUE),nacount)
```


================================
3. New Data after imputing missing value looks like:  

``` {r }
str ( impactData)
```

================================
4.Summarisizing new data and comparing with original:


``` {r NewMeanValue}

newdatestepsdata=tapply(impactData$steps,impactData$date,sum)
newdfdatestepsdata=as.data.frame(newdatestepsdata)
newdfdatestepsdata[,2]=unique(actData$date)
names(newdfdatestepsdata)=c("stepsmean","date")

qplot(newdfdatestepsdata[,1], xlab = expression(bold("Total number of Steps")),  geom = "bar",binwidth=1000)


newmeansteps=mean(newdfdatestepsdata[,1],na.rm=TRUE)
newmedsteps=median(newdfdatestepsdata[,1],na.rm=TRUE)
```

**We see very little difference in the estimates from earlier data set** 

New Mean of total number of steps taken per day is `r newmeansteps`  
Original Mean of total number of steps taken per day is `r meansteps`  


New Median of total number of steps taken per day is `r newmedsteps`  
Original  Median of total number of steps taken per day is `r medsteps`



## Are there differences in activity patterns between weekdays and weekends?

1. Creating a new factor variable to split the data as Weekend data and Weekday Data 

``` {r Daylist}
daylist=weekdays(as.POSIXlt(impactData$date))
ww=ifelse((daylist=="Sunday") | (daylist== "Saturday"),"Weekend","Weekday")

impactData["ww"]=as.factor(ww)
wwSplitData=split(impactData,impactData$ww)

weekendactData=wwSplitData$Weekend
weekdayactData=wwSplitData$Weekday

str(weekendactData)
str(weekdayactData)

```

==============================
2. Compare Weekend activity with Weekday activity



``` {r Panelplot}
## For weekend Data
WEintstepsdata=tapply(weekendactData$steps,weekendactData$interval,mean)
WEdfintstepsdata=as.data.frame(WEintstepsdata)
WEdfintstepsdata[,2]=unique(actData$interval)
WEdfintstepsdata[,3]=as.factor(rep("Weekend",nrow(WEdfintstepsdata)))
names(WEdfintstepsdata)=c("avgdailysteps","interval","WeekSplit")


## For Weekday Data
WDintstepsdata=tapply(weekdayactData$steps,weekdayactData$interval,mean)
WDdfintstepsdata=as.data.frame(WDintstepsdata)
WDdfintstepsdata[,2]=unique(actData$interval)
WDdfintstepsdata[,3]=as.factor(rep("Weekday",nrow(WDdfintstepsdata)))
names(WDdfintstepsdata)=c("avgdailysteps","interval","WeekSplit")

## Join the data
wwdfintstepsdata=rbind(WDdfintstepsdata,WEdfintstepsdata)


ggplot(wwdfintstepsdata, aes( interval,avgdailysteps)) + geom_line()+facet_grid(WeekSplit~.) +    xlab("5-minute Interval ") + ylab("Number of steps")


```